Metropolitan Areas (Deaths & Case)

### 2020-11-13 KPMG Team/Dashboard/visualization_scripts/msa_county_state_cities.Rmd
More Visualizations
Metropolitan Areas within States

# plotting 
library(ggplot2)
# library(hrbrthemes) 
# library(geofacet)

@@

# metropolitan state areas delineation from file taken off Census Bureau
data <- read_csv("data/msa.csv")

@@

# select only msa-related information 
data_msa <- data %>% 
  mutate(fips = paste(`FIPS State Code`, `FIPS County Code`, sep = "")) %>%
  mutate(fips_state = `FIPS State Code`) %>%
  mutate(state = `State Name`) %>%
  mutate(fips_county = `FIPS County Code`) %>%
  rename(msa = `Metropolitan Division Code`) %>%
  rename(msa_title = `Metropolitan Division Title`) %>% 
  select(msa, msa_title, fips, state, fips_state, fips_county)

@@

# only need county to merge both datasets, will keep just metro areas (csa)
msa_covid_merge <- census_msa %>% 
  left_join(covid, by = "fips") %>%  # join nytimes county-covid df with population and msa ares
  select(msa, msa_title, msa_population, date, cases, deaths) %>% 
  # set date range
  filter(date >= lubridate::as_date("2020-01-24") &
           date <= lubridate::as_date(Sys.Date() - 8)) # 8 days prior, so that the 7 day mean wouldn't get messed up

@@

@@

@@

# get lat long data from US Census 
# https://www.census.gov/geographies/mapping-files/time-series/geo/carto-boundary-file.html
county_lines <- st_read("county_shape/cb_2018_us_county_500k.shp")

# add FIPS as a variable from State Fips and County Fips
county_lines <- county_lines %>% 
  mutate(fips = paste(STATEFP, COUNTYFP, sep = ""))

# add lat long attributes to county dataset          
county_geo <- merge(county_lines, county_sum, by = "fips")
county_map <- county_geo %>% st_transform('+proj=longlat +datum=WGS84')

# remove non-mainland US parts 
outside_mainland <- c("02", "15", "72", "2", "3")
islands <- c("Virgin Islands", "Northern Mariana Islands")

county_map_US <- county_map %>% 
  filter(!STATEFP %in% outside_mainland) %>% 
  filter(!state %in% islands) # remove alaska and hawaii

@@

@@

# TO DO: 
# Facet it

@@

ggsave("county-map-cases.pdf")

```
### Metropolitan Cases

>>>>

# plotting 
library(ggplot2)
library(ggthemes)
library(extrafonts)
# library(hrbrthemes) 
# library(geofacet)

@@

# set working directory to source file location

# metropolitan state areas delineation from file taken off Census Bureau
msa <- read_csv("data/msa.csv")

@@

# select only msa-related information 
data_msa <- msa %>% 
  mutate(fips = paste(`FIPS State Code`, `FIPS County Code`, sep = "")) %>%
  mutate(fips_state = `FIPS State Code`) %>%
  mutate(state = `State Name`) %>%
  mutate(fips_county = `FIPS County Code`) %>%
  rename(msa = `Metropolitan Division Code`) %>%
  rename(msa_title = `Metropolitan Division Title`) %>% 
  select(msa, msa_title, fips, state, fips_state, fips_county)

@@

# only need county to merge both datasets, will keep just metro areas (csa)
msa_county_covid_merge <- census_msa %>% 
  left_join(covid, by = c("fips", "state")) %>%
  select(!B01003_001M) %>%
  drop_na(msa)

# ###########
# sum of county, showing msa as well 
msa_county_sum <- msa_county_covid_merge %>%
  filter(date == max(date))
# ###########

# join nytimes county-covid df with population and msa areas (remove county)
msa_covid_merge <- msa_county_covid_merge %>% 
  select(msa, msa_title, msa_population, date, cases, deaths) %>% 
  # set date range
  filter(date >= lubridate::as_date("2020-01-24") &
           date <= lubridate::as_date(Sys.Date() - 8)) # 8 days prior, so that the 7 day mean wouldn't get messed up

@@

head(msa_county_sum)

@@

# final count of msa cases, in order to get ranking
state_sum <- state_covid %>% 
  filter(date == max(state_covid$date)) %>% 
  select(state, date, cases_sum, deaths_sum) %>% 
  distinct()

# --------------------------------------------------------- #
# ----------- GETTING STATE & MSA TOGETHER ---------------- # 
# --------------------------------------------------------- #

# making both datasets bindable 
msa_covid_sum_merge <- msa_covid_sum %>% 
  rename(msa_state = "msa_title") %>% 
  ungroup() %>%
  select(msa_state, date, cases_sum, deaths_sum)

state_covid_sum_merge <- state_covid_sum %>% 
  rename(msa_state = "state") %>% 
  select(msa_state, date, cases_sum, deaths_sum)

# binding row  
msa_state_covid_sum <- msa_covid_sum_merge %>%
  bind_rows(state_covid_sum_merge)

@@

# get lat long data from US Census 
# https://www.census.gov/geographies/mapping-files/time-series/geo/carto-boundary-file.html
county_lines <- st_read("data/county_shape/cb_2018_us_county_500k.shp")
state_lines <- st_read("data/state_shape/cb_2018_us_state_500k.shp")

names(state_sum)
head(state_lines)

# merge county 
# add FIPS as a variable from State Fips and County Fips
county_lines <- county_lines %>% 
  mutate(fips = paste(STATEFP, COUNTYFP, sep = ""))
# add lat long attributes to county dataset          
county_geo <- merge(county_lines, county_sum, by = "fips")
county_map <- county_geo %>% st_transform('+proj=longlat +datum=WGS84')

# merge MSA
msa_geo <- merge(county_lines, msa_county_sum, by = "fips")
msa_map <- msa_geo %>% st_transform('+proj=longlat +datum=WGS84')

# merge states and state lines
state_lines <- state_lines %>% 
  rename(state = NAME)
state_geo <- merge(state_lines, state_sum, by = "state")
state_map <- state_geo %>% st_transform('+proj=longlat +datum=WGS84')

# remove non-mainland US parts 
# outside_mainland <- c("02", "15", "72", "2", "3")
outside_mainland <- c("Alaska", 
                      "Hawaii", 
                      "Puerto Rico", 
                      "Virgin Islands", 
                      "Northern Mariana Islands",
                      "Guam")

county_map_US <- county_map %>% 
  filter(!state %in% outside_mainland)

msa_map_US <- msa_map %>% 
  filter(!state %in% outside_mainland)

state_map_US <- state_map %>% 
  filter(!state %in% outside_mainland) %>% 
  rename(cases = cases_sum) %>% 
  rename(deaths = deaths_sum)

names(state_map_US)

@@

ggplot(data = msa_map_US) +
  geom_sf(aes(fill = cases)) + 
  scale_fill_viridis_c(option = "magma", direction = 1)

ggplot(data = state_map_US) +
  geom_sf(aes(fill = cases)) + 
  scale_fill_viridis_c(option = "magma", direction = -1) + 
  theme_void()

@@

# TO DO:

 fips_code <- fips(state = "New York") 
# Facet it

@@

# msa state map 
msa_state_map <- function(msa) {

  msa_map_state <- msa_map %>% 
  filter(msa == msa)

  ggplot(data = msa_map_state) +
    geom_sf(aes(fill = cases)) + 
    scale_fill_viridis_c(option = "magma", direction = -1) + 
    labs(title = msa_map_state$NAME)

msa_state_map(40484)

head(msa_county_sum)

msa_map_NWNJ <- msa_map %>% 
  filter(msa == 35084)

31084

msa_map_LA <- msa_map %>% 
  filter(msa == 41884)

msa_map_SF <- msa_map %>% 
  filter(msa == 31084)

msa_map_OC <- msa_map %>% 
  filter(msa == 11244)

msa_map_OAK <- msa_map %>% 
  filter(msa == 36084)

msa_map_RA <- msa_map %>% 
  filter(msa == 42034)

state_map_CA <- state_map_US %>% 
  filter(state == "California")

state_map_CA_region <- state_map_CA

# state_map_CA <- state_map_CA %>% 
#   mutate(region_cases = cases - msa_map_LA$cases
#          
state_map_CA_region$cases <- state_map_CA$cases -
  sum(msa_map_LA$cases) - 
  msa_map_SF$cases - 
  msa_map_OC$cases - 
  sum(msa_map_OAK$cases) - 
  msa_map_RA$cases

ggplot() +
    geom_sf(data = state_map_CA_region, aes(fill = cases)) + 
    geom_sf(data = msa_map_LA, aes(fill = cases)) + 
    geom_sf(data = msa_map_SF, aes(fill = cases)) + 
    geom_sf(data = msa_map_OC, aes(fill = cases)) + 
    geom_sf(data = msa_map_OAK, aes(fill = cases)) + 
    geom_sf(data = msa_map_RA, aes(fill = cases)) + 
    scale_fill_viridis_c(option = "magma", direction = -1) + 
    labs(title = state_map_CA$state,
         subtitle = "Metropolitan Statistical Areas") + 
    theme_void()

ggplot() +
    geom_sf(data = state_map_CA, aes(fill = cases)) + 
    scale_fill_viridis_c(option = "magma", direction = -1) + 
    labs(subtitle = state_map_CA$state) + 
    theme_void()

names(msa_map_NWNJ)

### TEXAS ###

state_map_TX <- state_map_US %>% 
  filter(state == "Texas")

texas_msa <- msa_county_sum %>% 
  select("state", "msa_title", "msa") %>%
  filter(state == "Texas") %>%
  distinct()

texas_msa_code <- pull(texas_msa, msa) # get msa codes into a vector

msa_map_TX <- msa_map %>% 
  filter(msa %in% texas_msa_code)

state_map_TX$region_cases <- state_map_TX$cases - sum(msa_map_TX$cases)

ggplot() +
    geom_sf(data = state_map_TX, aes(fill = region_cases)) +
    geom_sf(data = msa_map_TX, aes(fill = cases)) + 
    scale_fill_viridis_c(option = "magma", direction = -1) + 
    labs(title = state_map_TX$state,
         subtitle = "Metropolitan Statistical Areas") + 
    theme_void()

head(msa_map_TX)

### FUNCTION ###

msa_state_map_US <- function(state_name) {

  # e.g. filter to get state's map
  state_map <- state_map_US %>% 
    filter(state == state_name)

  state_msa <- msa_county_sum %>% 
    select("state", "msa_title", "msa") %>%
    filter(state == state_name) %>%
    distinct()

  # get msas from the state 
  state_msa_code <- pull(state_msa, msa) # get msa codes into a vector

  msa_map_state <- msa_map %>% 
    filter(msa %in% state_msa_code)

  # get regional numbers
  state_map$region_cases <- state_map$cases - sum(msa_map_state$cases)

  #   # get two-letter e.g. NY
  state_code <- state_map$STUSPS

  # df <- list()
  # 
  # df$state_map <- state_map 
  # df$msa_map_state <- msa_map_state

#   return(df)
# }  
  # plot
  ggplot() +
    geom_sf(data = state_map, aes(fill = region_cases)) +
    geom_sf(data = msa_map_state, aes(fill = cases)) + 
    scale_fill_viridis_c(option = "magma", direction = -1) + 
    labs(title = state_map$state,
         subtitle = "Metropolitan Statistical Areas") + 
    theme_void()

35614 # nyc is missing

df_ny <- msa_state_map_US("New York")

View(df_ny$msa_map_state)

msa_map %>% 
  filter(state == "New York")

IL_msa <- msa_state_map_US("Illinois")
CA_msa <- msa_state_map_US("California")
TX_msa <- msa_state_map_US("Texas")
NJ_msa <- msa_state_map_US("New Jersey")
MA_msa <- msa_state_map_US("Massachusetts")
FL_msa <- msa_state_map_US("Florida")

plot_grid(TX_msa, CA_msa, FL_msa, IL_msa, MA_msa, NJ_msa, ncol = 2)

# ggsave("metro-state.png")
```
### Metropolitan Cases
